---
output:
  pdf_document: default
  html_document: default
---

Dustin Brewer ; Brooke Peckenpaugh


#Set working directory, load packages

```{r}
setwd("~/GitHub/GroupProject")
rm(list=ls())

library(vegan)
library(psych)
library(ggplot2)
library(forcats)
library(dplyr)
library(corrplot)
library(indicspecies)


```


# SPECIES RICHNESS CODE, WITHOUT FORAGING GUILDS

#Load, manipulate data

```{r}
data.without.guilds <- read.csv("hf085-01-bird.csv", header = TRUE)

data.num <- data.without.guilds[ ,3:51]

```


#Write functions

```{r}
S.obs <- function(x = ""){
  rowSums(x > 0) * 1
}

S.chao2 <- function(site = "", SbyS = ""){
  SbyS = as.data.frame(SbyS)
  x = SbyS[site, ]
  SbyS.pa <- (SbyS > 0) * 1
  Q1 = sum(colSums(SbyS.pa) == 1)
  Q2 = sum(colSums(SbyS.pa) == 2)
  S.chao2 = S.obs(x) + (Q1^2)/(2*Q2)
  return(S.chao2)
}

sem <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}

```


#Determine observed and estimated species richness
```{r}
obs.rich <- S.obs(data.num)
obs.rich


est.rich <- S.chao2(1:40, data.num)
est.rich
```

#Make species richness plot
```{r}
data.rich <- cbind(data.without.guilds, est.rich)

m1 <- lm(est.rich ~ mortality.class, data = data.rich)
summary(m1)


mortality <- factor(data.rich$mortality.class, levels = c("Low", "Med", "Hi-A", "Hi-B"))
mortality

data.means <- tapply(data.rich$est.rich, mortality, mean)


data.sem <- tapply(data.rich$est.rich, mortality, sem)


bp1 <- barplot(data.means, ylim = c(0, round(max(data.rich$est.rich), digits = 0)),
               pch = 15, cex = 1.25, las = 1, cex.lab = 1.4, cex.axis = 1.25,
               xlab = "Mortality class",
               ylab = "# Species",
               names.arg = c("Low", "Medium", "High-A", "High-B"), main = "Estimated Richness")

arrows(x0 = bp1, y0 = data.means, y1 = data.means - data.sem, angle = 90,
       length = 0.1, lwd = 1)
arrows(x0 = bp1, y0 = data.means, y1 = data.means + data.sem, angle = 90,
       length = 0.1, lwd = 1)


```

#Run Anova
```{r}
fitanova <- aov(est.rich ~ mortality, data = data.rich)
fitanova

summary(fitanova)
TukeyHSD(fitanova)

par(mfrow = c(2,2), mar = c(5.1, 4.1, 4.1, 2.1))
plot(fitanova)

```







#Mean abundance

#Open data 
```{r}
MeanAbundMort <- read.csv("MortalityMeanAbund.csv")
```


#Reorder labels on x axis

```{r}
MeanAbundMort$Mortality <- fct_relevel(MeanAbundMort$Mortality, "Low", "Medium", "High-B", "High-A")

```


#Plot

```{r}
ggplot(MeanAbundMort, aes(x=Mortality, y = Abundance)) +
  geom_bar(position=position_dodge(), stat ="identity", colour = 'black') + geom_errorbar(aes(ymin=Abundance-S.E., ymax=Abundance+S.E.), width = .2, position = position_dodge(.9)) + ggtitle("Mean Abundance") 

```






#FORAGING GUILDS CODE

#Species Richness


#Open species richness foraging guilds file

Spec.Rich.Guilds <- read.csv("GuildsClusteredRich.csv")
```{r}
Spec.Rich.Guilds <- read.csv("GuildsClusteredRich.csv")
```


#Reorder labels on x axis
```{r}

Spec.Rich.Guilds$Mortality <- fct_relevel(Spec.Rich.Guilds$Mortality, "Low", "Medium", "High-B", "High-A")
```


#Plot

```{r}
ggplot(Spec.Rich.Guilds, aes(x=Mortality, y = Richness, fill = Guild)) +
  geom_bar(position=position_dodge(), stat ="identity", colour = 'black') + geom_errorbar(aes(ymin=Richness-S.E., ymax=Richness+S.E.), width = .2, position = position_dodge(.9)) + ggtitle("Observed Species Richness") 
```


# Mean Abundance (Guilds)

#Load Data
```{r}
Guilds <- read.csv("ForagingGuildsDataFile.csv")

str(Guilds)
summary(Guilds)
```


#Compute means for omnivores and insectivores in the different mortality classes

```{r}
Omnivores.Mean <- tapply(Guilds$Omn_16, Guilds$mortality.class, mean)
Omnivores.Mean

Insectivores.Mean <- tapply(Guilds$Ins_29, Guilds$mortality.class, mean)
Insectivores.Mean

```



#Compute standard errors for those means

```{r}
Omnivores.SE <- tapply(Guilds$Omn_16, Guilds$mortality.class, sem)
Omnivores.SE


Insectivores.SE <- tapply(Guilds$Ins_29, Guilds$mortality.class, sem)
Insectivores.SE
```


#Omnivore bar plot

```{r}
OmnivorePlot <- barplot(Omnivores.Mean, ylim = c(0, round(1.5*max(Omnivores.Mean), digits = 0)),
               pch = 15, cex = 1.25, las = 1, cex.lab = 1.4, cex.axis = 1.25,
               xlab = "mortality class",
               ylab = "mean site abundance",
               names.arg = c("low", "medium", "high A", "high B"))

arrows(x0 = OmnivorePlot, y0 = Omnivores.Mean, y1 = Omnivores.Mean - Omnivores.SE, angle = 90,
       length = 0.1, lwd = 1)
arrows(x0 = OmnivorePlot, y0 = Omnivores.Mean , y1 = Omnivores.Mean + Omnivores.SE, angle = 90,
       length = 0.1, lwd = 1)




```


#Insectivore bar plot

```{r}
InsectivorePlot <- barplot(Insectivores.Mean, ylim = c(0, round(1.5*max(Insectivores.Mean), digits = 0)),
                        pch = 15, cex = 1.25, las = 1, cex.lab = 1.4, cex.axis = 1.25,
                        xlab = "mortality class",
                        ylab = "mean site abundance",
                        names.arg = c("low", "medium", "high A", "high B"))

arrows(x0 = InsectivorePlot, y0 = Insectivores.Mean, y1 = Insectivores.Mean - Insectivores.SE, angle = 90,
       length = 0.1, lwd = 1)
arrows(x0 = InsectivorePlot, y0 = Insectivores.Mean , y1 = Insectivores.Mean + Insectivores.SE, angle = 90,
       length = 0.1, lwd = 1)

```



#Get omnivore and insectivore bars on same plot

  #Load excel file with means and S.E.s for both guilds at each mortality class
```{r}
ClusteredBar <- read.csv("GuildsClusteredBar.csv", header = TRUE)
ClusteredBar
str(ClusteredBar)

```



#Reorder x axis

```{r}
ClusteredBar$Mortality <- fct_relevel(ClusteredBar$Mortality, "Low", "Medium", "High-B", "High-A")


ggplot(ClusteredBar, aes(x=Mortality, y = Mean, fill = Guild)) +
  geom_bar(position=position_dodge(), stat ="identity", colour = 'black') + geom_errorbar(aes(ymin=Mean-S.E., ymax=Mean+S.E.), width = .2, position = position_dodge(.9)) + ggtitle("Bird Abundance") 

```


#Run Omnivore Anova

```{r}

OmnivoreAnova <- aov(Guilds$Omn_16~Guilds$mortality.class, data = Guilds)
OmnivoreAnova

summary(OmnivoreAnova)
TukeyHSD(OmnivoreAnova)
```



#Run Insectivore ANOVA

```{r}
InsectivoreAnova <- aov(Guilds$Ins_29~Guilds$mortality.class, data = Guilds)

summary(InsectivoreAnova)
TukeyHSD(InsectivoreAnova)



```


#PCoA

```{r}
bird <- read.csv("hf085-01-bird.csv", header = TRUE)
over <- read.csv("hf085-02-overstory.csv", header = TRUE)
under <- read.csv("hf085-03-understory.csv", header = TRUE)

bird.num <- bird[3:51]
over.num <- over[3:23]
under.num <- under[3:20]

# Sorenson & PCoA
bird.rm <- vegdist(bird.num, method = "bray", binary = TRUE)
bird.pcoa <- cmdscale(bird.rm, eig = TRUE, k = 3)

explainvar1 <- round(bird.pcoa$eig[1] / sum(bird.pcoa$eig), 3) * 100
explainvar2 <- round(bird.pcoa$eig[2] / sum(bird.pcoa$eig), 3) * 100
explainvar3 <- round(bird.pcoa$eig[3] / sum(bird.pcoa$eig), 3) * 100
sum.eig <- sum(explainvar1, explainvar2, explainvar3)

# Plotting PCoA
palette(c('lightblue1', 'lightblue3', 'gray50', 'darkolivegreen3'))
par(mar = c(5, 5, 1, 2) + 0.1)
plot(bird.pcoa$points[ ,1], bird.pcoa$points[ ,2], ylim = c(-0.3, 0.6),
     xlab = paste("PCoA 1 (", explainvar1, "%)", sep = ""),
     ylab = paste("PCoA 2 (", explainvar2, "%)", sep = ""),
     pch = 16, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)
points(bird.pcoa$points[ ,1], bird.pcoa$points[ ,2],
       pch = 19, cex = 3, bg = bird$mortality.class, col = bird$mortality.class)
legend("topleft", legend=c("Low", "Med", "Hi-A", "Hi-B"), fill=c('gray50', 'darkolivegreen3', 'lightblue1', 'lightblue3'))

```


#Phi coefficient of association

```{r}
mortality <- c(rep("Hi-A", 10), rep("Hi-B", 11), rep("Low", 11), rep("Med", 8))

bird.rel <- decostand(bird.num, method = "total")
phi <- multipatt(bird.rel, cluster = mortality, func = "r.g", control = how(nperm=10000))
summary(phi)
```


#Mantel test

```{r}
bird.dist <- vegdist(bird.num, method = "bray")
over.dist <- vegdist(over.num, method = "bray")

bird2.dist <- vegdist(bird.num[-29,], method = "bray")
under.dist <- vegdist(under.num[-29,], method = "bray")

mantel(bird.dist, over.dist)
mantel(bird2.dist, under.dist)
```



